NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-03 paper/Appendices/Appendix0_10.tex
results MLCA almost done

>>>>

\begin{verbatim}
Mplus VERSION 7.4
MUTHEN & MUTHEN
07/28/2018   9:55 AM

INPUT INSTRUCTIONS

  TITLE:    3-class at level 1 (CW), 3-classes at level 2 (CB) random effects model - non-pa
            ordered polytomous variables for carb intake at each time slot over four
            days of NDNS survey 2008/09 - 2015/16
            variable 0 = not eating
                     1 = eating & carb provided < 50% calorie
                     2 = eating & carb provided >= 50% calorie

  DATA:     File is H:\summer_project\Mplus\TimeSlots\NDNS_Tslots.dat;

  VARIABLE: NAMES = id id_dy Age Sex H6_9 H9_12 H12_14 H14_17 H17_20
                    H20_22 H22_6;

            USEVAR = H6_9 H9_12 H12_14 H14_17 H17_20
                    H20_22 H22_6;

            auxiliary = Age Sex;

            CATEGORICAL =  H6_9 H9_12 H12_14 H14_17 H17_20
                    H20_22 H22_6;

            CLUSTER = id;

            IDVARIABLE = id_dy;

            BETWEEN = CB;

            WITHIN = H6_9 H9_12 H12_14 H14_17 H17_20
                    H20_22 H22_6;

            CLASSES = CB(3) CW(3);

            MISSING are .;

  ANALYSIS:
  type = mixture twolevel;
  starts = 50 25;
  process = 8(starts);

  MODEL:
  %within%
  %overall%
  %between%
  %overall%
  CW ON CB;

  Savedata:
    file is H:\summer_project\Mplus\TimeSlots\Multilevel\NDNSslot_CW3CB3.txt;
    save is cprob;
    format is free;

3-class at level 1 (CW), 3-classes at level 2 (CB) random effects model - non-par
ordered polytomous variables for carb intake at each time slot over four
days of NDNS survey 2008/09 - 2015/16
variable 0 = not eating
1 = eating & carb provided < 50% calorie
2 = eating & carb provided >= 50% calorie

SUMMARY OF ANALYSIS

Number of groups                                                 1
Number of observations                                       24483

Number of dependent variables                                    7
Number of independent variables                                  0
Number of continuous latent variables                            0
Number of categorical latent variables                           2

Observed dependent variables

  Binary and ordered categorical (ordinal)
   H6_9        H9_12       H12_14      H14_17      H17_20      H20_22
   H22_6

Observed auxiliary variables
   AGE         SEX

Categorical latent variables
   CB          CW

Variables with special functions

  Cluster variable      ID
  ID variable           ID_DY

  Within variables
   H6_9        H9_12       H12_14      H14_17      H17_20      H20_22
   H22_6

Estimator                                                      MLR
Information matrix                                        OBSERVED
Optimization Specifications for the Quasi-Newton Algorithm for
Continuous Outcomes
  Maximum number of iterations                                 100
  Convergence criterion                                  0.100D-05
Optimization Specifications for the EM Algorithm
  Maximum number of iterations                                 500
  Convergence criteria
    Loglikelihood change                                 0.100D-02
    Relative loglikelihood change                        0.100D-05
    Derivative                                           0.100D-02
Optimization Specifications for the M step of the EM Algorithm for
Categorical Latent variables
  Number of M step iterations                                    1
  M step convergence criterion                           0.100D-02
  Basis for M step termination                           ITERATION
Optimization Specifications for the M step of the EM Algorithm for
Censored, Binary or Ordered Categorical (Ordinal), Unordered
Categorical (Nominal) and Count Outcomes
  Number of M step iterations                                    1
  M step convergence criterion                           0.100D-02
  Basis for M step termination                           ITERATION
  Maximum value for logit thresholds                            15
  Minimum value for logit thresholds                           -15
  Minimum expected cell size for chi-square              0.100D-01
Maximum number of iterations for H1                           2000
Convergence criterion for H1                             0.100D-03
Optimization algorithm                                         EMA
Integration Specifications
  Type                                                    STANDARD
  Number of integration points                                  15
  Dimensions of numerical integration                            0
  Adaptive quadrature                                           ON
Random Starts Specifications
  Number of initial stage random starts                         50
  Number of final stage optimizations                           25
  Number of initial stage iterations                            10
  Initial stage convergence criterion                    0.100D+01
  Random starts scale                                    0.500D+01
  Random seed for generating random starts                       0
Parameterization                                             LOGIT
Link                                                         LOGIT
Cholesky                                                       OFF

Input data file(s)
  H:\summer_project\Mplus\TimeSlots\NDNS_Tslots.dat
Input data format  FREE

SUMMARY OF DATA

     Number of missing data patterns             1
     Number of y missing data patterns           0
     Number of u missing data patterns           1
     Number of clusters                       6155

COVARIANCE COVERAGE OF DATA

Minimum covariance coverage value   0.100

UNIVARIATE PROPORTIONS AND COUNTS FOR CATEGORICAL VARIABLES

    H6_9
      Category 1    0.313         7655.000
      Category 2    0.184         4500.000
      Category 3    0.504        12328.000
    H9_12
      Category 1    0.222         5447.000
      Category 2    0.295         7227.000
      Category 3    0.482        11809.000
    H12_14
      Category 1    0.195         4783.000
      Category 2    0.454        11112.000
      Category 3    0.351         8588.000
    H14_17
      Category 1    0.283         6926.000
      Category 2    0.338         8277.000
      Category 3    0.379         9280.000
    H17_20
      Category 1    0.124         3043.000
      Category 2    0.582        14240.000
      Category 3    0.294         7200.000
    H20_22
      Category 1    0.356         8722.000
      Category 2    0.363         8898.000
      Category 3    0.280         6863.000
    H22_6
      Category 1    0.666        16295.000
      Category 2    0.169         4144.000
      Category 3    0.165         4044.000

RANDOM STARTS RESULTS RANKED FROM THE BEST TO THE WORST LOGLIKELIHOOD VALUES

Final stage loglikelihood values at local maxima, seeds, and initial stage start numbers:

         -166348.815  153942           31
         -166348.815  573096           20
         -166348.815  253358           2
         -166348.816  318230           46
         -166348.816  246261           38
         -166348.873  285380           1
         -166348.908  903420           5
         -166349.394  120506           45
         -166349.394  966014           37
         -166349.394  207896           25
         -166349.395  195873           6
         -166349.513  68985            17
         -166349.514  366706           29
         -166352.737  76974            16
         -166357.057  127215           9
         -166482.723  533738           11
         -166495.844  645664           39
         -166668.918  372176           23

THE BEST LOGLIKELIHOOD VALUE HAS BEEN REPLICATED.  RERUN WITH AT LEAST TWICE THE
RANDOM STARTS TO CHECK THAT THE BEST LOGLIKELIHOOD IS STILL OBTAINED AND REPLICATED.

THE MODEL ESTIMATION TERMINATED NORMALLY

MODEL FIT INFORMATION

Number of Free Parameters                      134

Loglikelihood

          H0 Value                     -166348.815
          H0 Scaling Correction Factor      1.8182
            for MLR

Information Criteria

          Akaike (AIC)                  332965.630
          Bayesian (BIC)                334051.799
          Sample-Size Adjusted BIC      333625.950
            (n* = (n + 2) / 24)

MODEL RESULTS USE THE LATENT CLASS VARIABLE ORDER

   CB  CW

  Latent Class Variable Patterns

         CB        CW
      Class     Class

         1         1
         1         2
         1         3
         2         1
         2         2
         2         3
         3         1
         3         2
         3         3

FINAL CLASS COUNTS AND PROPORTIONS FOR THE LATENT CLASS PATTERNS
BASED ON ESTIMATED POSTERIOR PROBABILITIES

  Latent Class
    Pattern

    1  1       4050.97975          0.16546
    1  2       1561.55249          0.06378
    1  3       1286.46696          0.05255
    2  1       2746.94031          0.11220
    2  2       3011.00217          0.12298
    2  3       1341.59686          0.05480
    3  1       2748.25320          0.11225
    3  2       4770.55950          0.19485
    3  3       2965.64876          0.12113

FINAL CLASS COUNTS AND PROPORTIONS FOR EACH LATENT CLASS VARIABLE
BASED ON ESTIMATED POSTERIOR PROBABILITIES

  Latent Class
    Variable    Class

    CB             1      6898.99902          0.28179
                   2      7099.53906          0.28998
                   3     10484.46094          0.42823
    CW             1      9546.17285          0.38991
                   2      9343.11426          0.38162
                   3      5593.71240          0.22847

FINAL CLASS COUNTS AND PROPORTIONS FOR THE LATENT CLASS PATTERNS
BASED ON THEIR MOST LIKELY LATENT CLASS PATTERN

Class Counts and Proportions

  Latent Class
    Pattern

    1  1             4262          0.17408
    1  2             1406          0.05743
    1  3             1178          0.04812
    2  1             2807          0.11465
    2  2             2946          0.12033
    2  3             1260          0.05146
    3  1             2745          0.11212
    3  2             5315          0.21709
    3  3             2564          0.10473

FINAL CLASS COUNTS AND PROPORTIONS FOR EACH LATENT CLASS VARIABLE
BASED ON THEIR MOST LIKELY LATENT CLASS PATTERN

  Latent Class
    Variable    Class

    CB             1            6846          0.27962
                   2            7013          0.28644
                   3           10624          0.43393
    CW             1            9814          0.40085
                   2            9667          0.39485
                   3            5002          0.20431

CLASSIFICATION QUALITY

     Entropy                         0.630

Average Latent Class Probabilities for Most Likely Latent Class Pattern (Row)
by Latent Class Pattern (Column)

  Latent Class Variable Patterns

  Latent Class         CB        CW
   Pattern No.      Class     Class

         1             1         1
         2             1         2
         3             1         3
         4             2         1
         5             2         2
         6             2         3
         7             3         1
         8             3         2
         9             3         3

           1        2        3        4        5        6        7        8        9

    1   0.720    0.091    0.073    0.016    0.032    0.004    0.005    0.033    0.025
    2   0.183    0.609    0.098    0.005    0.002    0.030    0.040    0.005    0.027
    3   0.211    0.084    0.629    0.008    0.005    0.007    0.011    0.036    0.009
    4   0.019    0.004    0.002    0.692    0.184    0.051    0.011    0.034    0.003
    5   0.042    0.001    0.001    0.158    0.709    0.045    0.001    0.035    0.009
    6   0.012    0.037    0.013    0.065    0.084    0.702    0.042    0.003    0.042
    7   0.011    0.029    0.004    0.012    0.002    0.022    0.641    0.126    0.153
    8   0.026    0.003    0.009    0.025    0.024    0.001    0.115    0.675    0.123
    9   0.046    0.024    0.004    0.003    0.010    0.018    0.079    0.174    0.642

MODEL RESULTS

                                                    Two-Tailed
                    Estimate       S.E.  Est./S.E.    P-Value

Within Level

Latent Class Pattern 1 1

 Thresholds
    H6_9$1            -0.718      0.218     -3.294      0.001
    H6_9$2             0.973      0.299      3.258      0.001
    H9_12$1           -2.516      0.463     -5.433      0.000
    H9_12$2            0.675      0.132      5.118      0.000
    H12_14$1          -1.025      0.145     -7.057      0.000
    H12_14$2           1.240      0.116     10.725      0.000
    H14_17$1          -1.566      0.149    -10.520      0.000
    H14_17$2           1.090      0.100     10.909      0.000
    H17_20$1          -1.998      0.125    -16.000      0.000
    H17_20$2           1.549      0.100     15.556      0.000
    H20_22$1          -0.933      0.085    -10.914      0.000
    H20_22$2           1.829      0.103     17.770      0.000
    H22_6$1            0.253      0.083      3.046      0.002
    H22_6$2            2.308      0.117     19.691      0.000

Latent Class Pattern 1 2

 Thresholds
    H6_9$1            -4.021      1.788     -2.249      0.025
    H6_9$2            -0.115      0.259     -0.445      0.656
    H9_12$1            0.167      0.373      0.448      0.654
    H9_12$2            2.142      0.586      3.657      0.000
    H12_14$1          -3.210      1.518     -2.115      0.034
    H12_14$2           0.858      0.167      5.124      0.000
    H14_17$1           0.044      0.384      0.114      0.909
    H14_17$2           1.617      0.293      5.509      0.000
    H17_20$1          -2.109      0.390     -5.409      0.000
    H17_20$2           1.399      0.196      7.126      0.000
    H20_22$1          -0.367      0.174     -2.109      0.035
    H20_22$2           2.347      0.382      6.151      0.000
    H22_6$1            0.754      0.259      2.912      0.004
    H22_6$2            2.542      0.264      9.646      0.000

Latent Class Pattern 1 3

 Thresholds
    H6_9$1           -15.000      0.000    999.000    999.000
    H6_9$2             2.357      0.783      3.011      0.003
    H9_12$1           -1.433      0.372     -3.850      0.000
    H9_12$2           -0.604      0.279     -2.166      0.030
    H12_14$1          -1.988      0.257     -7.749      0.000
    H12_14$2           0.524      0.125      4.209      0.000
    H14_17$1          -1.027      0.232     -4.436      0.000
    H14_17$2           0.274      0.131      2.087      0.037
    H17_20$1          -2.665      0.310     -8.605      0.000
    H17_20$2           0.707      0.112      6.322      0.000
    H20_22$1          -0.527      0.152     -3.462      0.001
    H20_22$2           0.702      0.138      5.102      0.000
    H22_6$1            1.119      0.185      6.062      0.000
    H22_6$2            1.748      0.183      9.544      0.000

Latent Class Pattern 2 1

 Thresholds
    H6_9$1             1.663      0.199      8.370      0.000
    H6_9$2             1.839      0.198      9.274      0.000
    H9_12$1           -2.150      0.281     -7.643      0.000
    H9_12$2           -0.869      0.140     -6.190      0.000
    H12_14$1          -1.978      0.191    -10.349      0.000
    H12_14$2           0.323      0.078      4.139      0.000
    H14_17$1           0.237      0.183      1.293      0.196
    H14_17$2           0.782      0.123      6.352      0.000
    H17_20$1          -2.936      0.428     -6.853      0.000
    H17_20$2           0.632      0.081      7.807      0.000
    H20_22$1           0.028      0.142      0.194      0.846
    H20_22$2           0.868      0.086     10.145      0.000
    H22_6$1            0.658      0.109      6.010      0.000
    H22_6$2            1.326      0.100     13.215      0.000

Latent Class Pattern 2 2

 Thresholds
    H6_9$1             1.640      0.171      9.619      0.000
    H6_9$2             1.906      0.179     10.678      0.000
    H9_12$1           -1.954      0.347     -5.636      0.000
    H9_12$2           -0.360      0.127     -2.842      0.004
    H12_14$1          -0.016      0.189     -0.084      0.933
    H12_14$2           0.948      0.135      7.029      0.000
    H14_17$1          -1.906      0.301     -6.327      0.000
    H14_17$2           0.371      0.080      4.614      0.000
    H17_20$1          -0.812      0.116     -7.030      0.000
    H17_20$2           0.910      0.089     10.259      0.000
    H20_22$1          -0.742      0.089     -8.318      0.000
    H20_22$2           0.998      0.085     11.705      0.000
    H22_6$1            0.298      0.083      3.608      0.000
    H22_6$2            1.337      0.099     13.475      0.000

Latent Class Pattern 2 3

 Thresholds
    H6_9$1            -1.072      0.500     -2.144      0.032
    H6_9$2            -0.309      0.346     -0.892      0.372
    H9_12$1            2.441      1.044      2.339      0.019
    H9_12$2            3.599      1.983      1.815      0.069
    H12_14$1          -1.029      0.211     -4.880      0.000
    H12_14$2           0.603      0.123      4.913      0.000
    H14_17$1          -0.010      0.243     -0.041      0.967
    H14_17$2           0.784      0.157      4.977      0.000
    H17_20$1          -0.953      0.203     -4.684      0.000
    H17_20$2           0.779      0.135      5.784      0.000
    H20_22$1          -0.105      0.210     -0.500      0.617
    H20_22$2           1.203      0.135      8.914      0.000
    H22_6$1            0.582      0.299      1.950      0.051
    H22_6$2            1.370      0.206      6.653      0.000

Latent Class Pattern 3 1

 Thresholds
    H6_9$1            -4.593      1.699     -2.703      0.007
    H6_9$2            -2.975      0.428     -6.957      0.000
    H9_12$1           -0.322      0.207     -1.553      0.120
    H9_12$2            0.398      0.363      1.095      0.274
    H12_14$1          -5.060      3.668     -1.380      0.168
    H12_14$2           0.307      0.100      3.080      0.002
    H14_17$1           0.186      0.530      0.351      0.726
    H14_17$2           0.317      0.245      1.295      0.195
    H17_20$1          -4.019      0.957     -4.199      0.000
    H17_20$2           0.747      0.093      7.987      0.000
    H20_22$1          -0.233      0.132     -1.767      0.077
    H20_22$2           0.607      0.109      5.571      0.000
    H22_6$1            1.304      0.146      8.918      0.000
    H22_6$2            1.850      0.160     11.579      0.000

Latent Class Pattern 3 2

 Thresholds
    H6_9$1            -1.232      0.195     -6.305      0.000
    H6_9$2            -0.858      0.169     -5.068      0.000
    H9_12$1           -4.377      1.937     -2.260      0.024
    H9_12$2           -1.488      0.316     -4.717      0.000
    H12_14$1          -1.727      0.227     -7.611      0.000
    H12_14$2           0.302      0.082      3.666      0.000
    H14_17$1          -1.834      0.237     -7.730      0.000
    H14_17$2          -0.294      0.186     -1.582      0.114
    H17_20$1          -2.588      0.487     -5.313      0.000
    H17_20$2           0.631      0.062     10.187      0.000
    H20_22$1          -0.920      0.078    -11.852      0.000
    H20_22$2           0.462      0.073      6.308      0.000
    H22_6$1            0.640      0.119      5.361      0.000
    H22_6$2            1.162      0.129      9.039      0.000

Latent Class Pattern 3 3

 Thresholds
    H6_9$1            -4.941      5.813     -0.850      0.395
    H6_9$2            -2.680      0.887     -3.024      0.002
    H9_12$1           -0.765      0.640     -1.195      0.232
    H9_12$2            1.164      0.920      1.265      0.206
    H12_14$1          -1.415      0.439     -3.226      0.001
    H12_14$2           0.566      0.085      6.626      0.000
    H14_17$1          -2.052      0.650     -3.158      0.002
    H14_17$2           0.612      0.210      2.909      0.004
    H17_20$1          -1.627      0.427     -3.810      0.000
    H17_20$2           0.713      0.103      6.935      0.000
    H20_22$1          -0.850      0.329     -2.585      0.010
    H20_22$2           0.685      0.134      5.104      0.000
    H22_6$1            1.237      0.195      6.349      0.000
    H22_6$2            1.893      0.179     10.582      0.000

Between Level

Categorical Latent Variables

Within Level

 Intercepts
    CW#1              -0.076      0.366     -0.208      0.835
    CW#2               0.475      0.309      1.539      0.124

Between Level

 CW#1       ON
    CB#1               1.223      0.473      2.585      0.010
    CB#2               0.793      0.441      1.796      0.073

 CW#2       ON
    CB#1              -0.282      0.535     -0.526      0.599
    CB#2               0.333      0.455      0.733      0.464

 Means
    CB#1              -0.417      0.100     -4.178      0.000
    CB#2              -0.386      0.067     -5.770      0.000

QUALITY OF NUMERICAL RESULTS

     Condition Number for the Information Matrix              0.428E-04
       (ratio of smallest to largest eigenvalue)

SAVEDATA INFORMATION

  Save file
    H:\summer_project\Mplus\TimeSlots\Multilevel\NDNSslot_CW3CB3.txt

  Order of variables

    H6_9
    H9_12
    H12_14
    H14_17
    H17_20
    H20_22
    H22_6
    ID_DY
    AGE
    SEX
    CPROB1
    CPROB2
    CPROB3
    CPROB4
    CPROB5
    CPROB6
    CPROB7
    CPROB8
    CPROB9
    CB
    CW
    MLCJOINT
    ID

  Save file format           Free

  Save file record length    10000

DIAGRAM INFORMATION

  Mplus diagrams are currently not available for Mixture analysis.
  No diagram output was produced.

     Beginning Time:  09:55:10
        Ending Time:  10:02:01
       Elapsed Time:  00:06:51

MUTHEN & MUTHEN
3463 Stoner Ave.
Los Angeles, CA  90066

Tel: (310) 391-9971
Fax: (310) 391-8971
Web: www.StatModel.com
Support: Support@StatModel.com

Copyright (c) 1998-2015 Muthen & Muthen
\end{verbatim}

### 2018-08-03 paper/Chapters/Chapter2.tex
results MLCA almost done

Multilevel latent class analysis accounts for the nested structure of the data by allowing latent class intercepts to vary across level 2 units and thereby examining if and how level 2 units influence the level 1 latent classes. These random intercepts allow the probability of membership in a particular level 1 (observation days) latent class to vary across level 2 units (e.g., here in the current context are the individuals). Essentially this allows the probability that an observation day will belong to a
particular day-level latent class to vary across individual levels.\vspace{-0.5cm}

@@

Proposed by Vermunt \parencite{Vermunt, vermunt2008latent} and Asparouhov and Muth\'en \parencite{muthen2009multilevel},  a traditional, parametric approach can be applied using a logistic regression model. For example, let's assume that there are two types of observation days in the dietary survey--high and low carbohydrate eating days. In an unconditional logistic regression model, the probability of the outcome (i.e. an observed high carbohydrate eating day vs. a low carbohydrate eating day) is constant within individual-level which means for each person throughout his/her survey there is some probability of following a high carbohydrate eating day. A random effect model considers the individuals (level 2) to be drawn from the adult population in the UK, and the probability of the outcome (i.e. high carbohydrate eating days) across individuals is considered to be a random variable \parencite{snijders2011multilevel}.

@@

Same framework can be used to consider random effects in an LCA model, but instead of saying that $C_{ij}$ is either low or high carbohydrate eating days as if we already know, it is now replaced by a latent variable $G_{ij}$ which indicates the typologies of carbohydrate eating patterns. Then we can use the day-level data to assess the log-odds of belonging to $k$th type of carbohydrate eating pattern on a specific day of survey, and we allow the log-odds to vary across individuals. Therefore, for some persons the log-odds of having a $k$th type of carbohydrate eating pattern during the survey can be high, but for the other persons, the log-odds of following the $k$th type of carbohydrate eating pattern can be low.

If the day-level LCA model (carbohydrate eating pattern typologies) is best defined by $T (T \geqslant 2)$ latent classes, then $T-1$ random intercept will be specified by a two-level multinomial logistic regression model. Similar to the typical LCA models, the latent class variable in a MLCA is defined by multiple observed indicators (here is defined by the responses of eating carbohydrate within each time slots, throughout 4 consecutive days of survey period). Considering the latent class indicators are indicator variables ($U_{ijk}$), the MLCA model can be written as follows:\vspace{-1cm}

@@

\begin{itemize}
	\item $ U_{ijk} $ represents the response of eating carbohydrate (one of the following: not eating any food, $< 50\%$ of the energy, or $\geqslant 50\%$ of the energy) on $i$th day of the survey ($i \in (1,2,3,4)$) in $j$th individual at the $k$th time slot of the day ($k \in (1, 2, 3, \cdots, 7)$);
	\item $G_{ij}$ denotes the latent class membership for $j$th individuals on $i$th day of the survey, the total number of day-level latent class is $T$;
%	\item A specific latent class is referred to as $t$, and the total number of level 1 latent classes is denoted by $T$;
	\item $P(U_{ijk} = s_k|C_{ij} = t)$ is the probability of a specific response pattern, conditional on membership in latent class $t$.
\end{itemize}

@@

Since the parametric approach discussed above can be extremely computationally demanding \parencite{van2008using, vermunt2008latent}, an alternative approach is using a non-parametric MLCA \parencite{davidian2008growth}. In this approach, separate latent class models are specified for level 1 (observation days) and level 2 (individuals). Similar with the parametric MLCA approach, there are $T-1$ random intercepts, where $T$ is the number of level 1 latent classes. However, rather than assuming the random intercepts following a normal distribution, the non-parametric MLCA assumes a multinomial (discrete) distribution of the level 2 latent classes. This approach is less computationally demanding compared with the parametric approach. These level 2 (individual) latent classes reflect differences in the probability of belonging to a specific day-level latent class, so that individuals that contain observation days with similar probabilities for the level 1 latent classes will be grouped together. The non-parametric MLCA model can be defined as follows: \vspace{-0.7cm}

@@

\begin{itemize}
	\item $CB_j$ is individual-level latent class membership for $j$th individual;
	\item $\gamma_{tm}$ is day-level and individual-level indicators. 
\end{itemize}

@@

\begin{itemize}
	\item Firstly, we ignored the multilevel structure of the data and estimated a series of traditional LC models to determine the number of classes at the observational-level;
	\item Next, a series of MLCA models were fitted to account for the multi-level structrure of the data. In these models, the number of observational-level classes was based on the best fitting LCA model from the first step, and the LCA model at the individual-level was estimated to identify the number of individual-level latent classes;
	\item Thirdly, when number of individual-level latent classes is defined based on the previous stage, observational-level classes was modified (one class lower and one class higher than in the second step), to see the effect of changing level 1 classes and confirm the best fitting model.
\end{itemize}

The number of classes in either level 1 were determined by 1) the evaluation of model fit indices, including the Akaike information criterion (AIC), Bayesian information criterion (BIC), adjusted Bayesian information criterion (aBIC) where smaller values indicate better, and entropy which is a statistic that summarizes latent class probabilities where values near 1 indicate better latent class separation; 2) the Lo-Mendell-Rubin Likelihood Ratio Test (LMR-LRT) \parencite{lo2001testing, nylund2007deciding} which compare $q$ vs. $q-1$ class models, where $q$ is the number of latent classes and 3) pattern interpretability. In the step of performing multilevel LCA, where LMR-LRT is available, same rules of model fit indices and pattern interpretability were used to determine the optimal combination of latent classes in observation day-level and individual-level. MLCA models were fitted in Mplus 7.4 \parencite{muthen2005mplus}, the Mplus codes are shown in the Appendix xxx. \vspace{-0.3cm}

@@

\section{Association between day-level and individual-level latent classes and other variables}\vspace{-0.3cm}

Day-level latent classes identified by MLCA steps were tabulated according to whether the diary was recorded during weekdays or not. A contingency table giving the frequency of responses across the 7 time slots of the survey days was produced.

>>>>

Multilevel latent class analysis accounts for the nested structure of the data by allowing latent class intercepts to vary across level 2 units and thereby examining if and how level 2 units influence the level 1 latent classes. These random intercepts allow the probability of membership in a particular level 1 (observation days) latent class to vary across level 2 units (e.g., here in the current context are the individuals). Essentially this allows the probability that an observation day will belong to a
particular day level latent class to vary across individual levels.\vspace{-0.5cm}

@@

Proposed by Vermunt \parencite{Vermunt, vermunt2008latent} and Asparouhov and Muth\'en \parencite{muthen2009multilevel},  a traditional, parametric approach can be applied using a logistic regression model. For example, let's assume that there are two types of observation days in the dietary survey--high and low carbohydrate eating days. In an unconditional logistic regression model, the probability of the outcome (i.e. an observed high carbohydrate eating day vs. a low carbohydrate eating day) is constant within individual level which means for each person throughout his/her survey there is some probability of following a high carbohydrate eating day. A random effect model considers the individuals (level 2) to be drawn from the adult population in the UK, and the probability of the outcome (i.e. high carbohydrate eating days) across individuals is considered to be a random variable \parencite{snijders2011multilevel}.

@@

Same framework can be used to consider random effects in an LCA model, but instead of saying that $C_{ij}$ is either low or high carbohydrate eating days as if we already know, it is now replaced by a latent variable $G_{ij}$ which indicates the typologies of carbohydrate eating patterns. Then we can use the day level data to assess the log-odds of belonging to $k$th type of carbohydrate eating pattern on a specific day of survey, and we allow the log-odds to vary across individuals. Therefore, for some persons the log-odds of having a $k$th type of carbohydrate eating pattern during the survey can be high, but for the other persons, the log-odds of following the $k$th type of carbohydrate eating pattern can be low.

If the day level LCA model (carbohydrate eating pattern typologies) is best defined by $T (T \geqslant 2)$ latent classes, then $T-1$ random intercept will be specified by a two-level multinomial logistic regression model. Similar to the typical LCA models, the latent class variable in a MLCA is defined by multiple observed indicators (here is defined by the responses of eating carbohydrate within each time slots, throughout 4 consecutive days of survey period). Considering the latent class indicators are indicator variables ($U_{ijk}$), the MLCA model can be written as follows:\vspace{-1cm}

@@

\begin{itemize}
	\item $ U_{ijk} $ represents the response of eating carbohydrate (one of the following: not eating any food, $< 50\%$ of the energy, or $\geqslant 50\%$ of the energy) on $i$th day of the survey ($i \in (1,2,3,4)$) in $j$th individual at the $k$th time slot of the day ($k \in (1, 2, 3, \cdots, 7)$);
	\item $G_{ij}$ denotes the latent class membership for $j$th individuals on $i$th day of the survey, the total number of day level latent class is $T$;
%	\item A specific latent class is referred to as $t$, and the total number of level 1 latent classes is denoted by $T$;
	\item $P(U_{ijk} = s_k|C_{ij} = t)$ is the probability of a specific response pattern, conditional on membership in latent class $t$.
\end{itemize}

@@

Since the parametric approach discussed above can be extremely computationally demanding \parencite{van2008using, vermunt2008latent}, an alternative approach is using a non-parametric MLCA \parencite{davidian2008growth}. In this approach, separate latent class models are specified for level 1 (observation days) and level 2 (individuals). Similar with the parametric MLCA approach, there are $T-1$ random intercepts, where $T$ is the number of level 1 latent classes. However, rather than assuming the random intercepts following a normal distribution, the non-parametric MLCA assumes a multinomial (discrete) distribution of the level 2 latent classes. This approach is less computationally demanding compared with the parametric approach. These level 2 (individual) latent classes reflect differences in the probability of belonging to a specific day level latent class, so that individuals that contain observation days with similar probabilities for the level 1 latent classes will be grouped together. The non-parametric MLCA model can be defined as follows: \vspace{-0.7cm}

@@

\begin{itemize}
	\item $CB_j$ is individual level latent class membership for $j$th individual;
	\item $\gamma_{tm}$ is day level and individual level indicators. 
\end{itemize}

@@

\begin{itemize}
	\item Firstly, we ignored the multilevel structure of the data and estimated a series of traditional LC models to determine the number of classes at the observational-level;
	\item Next, a series of MLCA models were fitted to account for the multi-level structrure of the data. In these models, the number of observational-level classes was based on the best fitting LCA model from the first step, and the LCA model at the individual level was estimated to identify the number of individual level latent classes;
	\item Thirdly, when number of individual level latent classes is defined based on the previous stage, observational-level classes was modified (one class lower and one class higher than in the second step), to see the effect of changing level 1 classes and confirm the best fitting model.
\end{itemize}

The number of classes in either level 1 were determined by 1) the evaluation of model fit indices, including the Akaike information criterion (AIC), Bayesian information criterion (BIC), adjusted Bayesian information criterion (aBIC) where smaller values indicate better, and entropy which is a statistic that summarizes latent class probabilities where values near 1 indicate better latent class separation; 2) the Lo-Mendell-Rubin Likelihood Ratio Test (LMR-LRT) \parencite{lo2001testing, nylund2007deciding} which compare $q$ vs. $q-1$ class models, where $q$ is the number of latent classes and 3) pattern interpretability. In the step of performing multilevel LCA, where LMR-LRT is available, same rules of model fit indices and pattern interpretability were used to determine the optimal combination of latent classes in observation day level and individual level. MLCA models were fitted in Mplus 7.4 \parencite{muthen2005mplus}, the Mplus codes are shown in \textbf{Appendix \ref{AppendixB}}. \vspace{-0.3cm}

@@

\section{Day level latent classes and their characteristics}\vspace{-0.3cm}

Day level latent classes identified by MLCA steps were tabulated according to whether the diary was recorded during weekdays or not. A contingency table giving the frequency of responses across the 7 time slots of the survey days was produced.
